Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2023-02-23 paper/paper.tex
Revise section 7

The main organizing concept is that of an architecture-to-architecture \emph{security-improving transform}, achieved via the insertion of new architectural components aimed at mitigating cyber-vulnerabilities.
%
We describe two such transformations in this paper: (1) the insertion of a filter to prevent malformed data from a malicious actor being propagated to downstream components, and (2) the insertion of a runtime monitor to detect (and alert) unexpected behaviors arising from untrusted components.
%
We refer to these as \emph{high-assurance components} because they must be correct and fulfill their intended purpose for the system to be cyber-resilient.

@@

We report on a case study applying these transformations with \brfcs\ to an Unmanned Aerial Vehicle (UAV) system that uses the Air Force Research Laboratory's OpenUxAS services for route planning.
%
OpenUxAS, as an open source product, is considered \emph{untrusted}.
%
The UAV system is thus transformed to be resilient to malicious behavior that may arise in the untrusted component.
%
In particular, filters are added to guard against malformed input and a monitor is created to guard against malicious flight plans generated from OpenUxAS.
%
The case study system is complex and shows the viability of the approach in potential full-scale industrial design.

@@

Having passed \agr\ verification, the workflow in \figref{fig:workflow} proceeds to the thread analysis as the UAV must actively protect itself from malicious actors.
%
Threat analysis identifies the potential of a supply chain attack through the AI route planner since its third-party.
%
To mitigate the identified supply chain attack, the designer adds a requirement that the system must guard against malicious AI behavior.
%
This requirement is reflected in the component contracts by removing all output guarantees from the AI contract since that component is considered untrusted.
%
Without output guarantees, the AI component is able to output arbitrary events in the \agr\ verification analysis.

@@

As shown in \figref{fig:hardened-certificate}, the cyber-hardened system passes AGREE verification.
%
That means that the high-assurance components protect the WM from the untrusted AI generating malformed data, spontaneous responses, or delayed responses.
%
Synthesis must generate faithful implementations of the code contracts for the \agr\ verification result to apply to the deployed system.

@@

\item the wellformedness predicate is evaluated on the input;

@@

In this section, we outline the application of the \brfcs\ tool towards the development of a UAV surveillance system.
%
The motivation was to showcase the utility and capability of the BriefCASE toolchain on an industrial scale problem, to determine whether the technology provides designers with early validation of their model;
%
produces early information about cyber-vulnerabilities;
%
provides effective mitigations;
%
and handles a complex real-world message format, namely the UxAS message format.

@@

For this application, UxAS accepts three types of messages.
%
The \textit{Operating Region} message defines where the UAV can and cannot fly.
%
The \textit{Line Search Task} message contains a series of waypoints that the UAV should traverse.
%
The waypoints typically lie along some geographical feature of interest, such as a river or railway.
%
Note that the UAV may not be able to directly traverse the Line Search Task waypoints due to no-fly zone constraints specified in the Operating Region message.
%
Anytime after receiving the Operating Region and Line Search Task messages, a Ground Station can transmit an \textit{Automation Request} message, which instructs UxAS to generate a flight plan that satisfies these constraints.
%
UxAS passes the flight plan in an \textit{Automation Response} message to the Waypoint Manager.
%
Because the Flight Control Computer can only process a small number of waypoints at a time, the Waypoint Manager parcels a small number of waypoints corresponding to the current UAV position, and sends them to the Flight Control Computer over a serial connection via the UART Driver.

@@

In \brfcs, we analyze the model using one (or more) of the integrated cybersecurity analysis tools.
%
As the designers of the system, we reviewed the list of potential vulnerabilities, and then added new requirements to the system model.
%
For example, the vulnerbility analysis noted that UxAS is an open source project, so we marked that component as \texttt{uncontroled} (colored red in \figref{fig:sw-initial}) and added requirements for ensuring that unverified or malicious code (which could potentially be embedded in the component) is not able impact other components.

In total, we added seven cyber requirements from the vulnerability analysis into our model.
%
These include four wellformedness requirements on data, two requirements for monitoring the behavior of the open-source UxAS component, and an attestation requirement for ensuring the integrity of the Ground Station software.
%
Requirements are imported into the model as goals in the assurance case.
%
Because we can build the assurance case at any time during development, we can easily determine for a given snapshot of the model which requirements are not yet supported by evidence.

The intent of the wellformedness requirements is to prevent malformed messages from causing a buffer overrun or code injection attack.
%
In the UAV design, such messages are most likely to originate from a remote source or the uncontrolled UxAS component.
%
By placing filters on the connections upstream of mission-critical components, such attacks could be mitigated.
%
The Filter transform is therefore applied for each wellformedness requirement, inserting filter components on the incoming and outgoing UxAS connections.

The filter code contracts for the four UxAS connections are similar, and check that record values contained in the messages are within appropriate ranges.
%
For example, the Automation Response message filter, which drops messages containing malformed flight plans, is defined as shown in \figref{fig:automation-response-filter}.
%
Although \texttt{Latitude}, \texttt{Longitude} and \texttt{Altitude} are defined as 64-bit floating-point values, the filter only passes messages containing waypoint values between [-90,90], [-180,180], and [0,15000], respectively.

@@

In addition to monitoring the UxAS output for malformed messages, we must also monitor for suspicious behavior.
%
This requires adding components for detecting that UxAS has crashed, as well as monitoring the correctness of the flight plans it produces.
%
The Monitor transform is applied for this class of mitigation.
%
We add a \textit{response} monitor to send an alert if UxAS does not emit a response within a set amount of time from receiving a request.
%
And we add a \textit{geofence} monitor to ensure that generated flight plans are compliant with the specified keep-in and keep-out zones.
%
The Geofence Monitor is a gated monitor in that it prevents the observed automation response from reaching the Waypoint Manager if the zones are violated.
%
The Geofence Monitor code contract is shown in \figref{fig:geofence-monitor}.

@@

So far, we have addressed requirements that mitigate vulnerabilities related to malformed messages and malicious behavior \textit{on-board} the UAV.
%
But we also want to protect against a compromised ground station transmitting well-formed, but malicious, commands.
%
The final cyber requirement is mitigated by a transform that adds two components to the UAV software: an Attestation Manager for evaluating remote systems like the Ground Station, and an Attestation Gate for filtering messages from sources that have not been approved by the Attestation Manager.
%
The Attestation Manager \ckml\ implementation is automatically inserted into the application code base by \brfcs.
%
The code contract for the Attestation Gate is completely generated by the transform since its behavior is specific to attestation.

After transforming the model to address the cyber requirements, the software architecture now appears as shown in \figref{fig:hardened-sw}.
%
The shaded components (green) were added by the transform with the exception of the UxAS component (red).

After unit testing the monitors, we formally verify the model with \agr\ to show that all of the component contracts are satisfied.
%
We then run \splt\ to produce provably-correct code.
%
The synthesized code is output to a directory in the build file system with the location specified for each component in the model.
%
The corresponding correctness proof is used in our assurance case as additional evidence that the vulnerability has been properly mitigated.

@@

Once we have determined that the model is correct and satisfies its cyber requirements, and the software components in the model have been implemented, either by \splt\ or other means, the system can be built and deployed.
%
We run the HAMR tool to generate the component stubs and infrastructure code necessary to enable component communication and execution according to a specified schedule.
%
HAMR then compiles the software to execute on seL4.

The UAV and Ground Station software were deployed on ODROID XU4 hardware, and communicated with each other over ethernet.
%
The AMASE flight simulator, representing the Flight Control Computer was run on a Linux machine and connected to the UAV ODROID via a serial connection.
%
The UxAS implementation on the UAV was modified by adding malicious code that would prevent it from responding to Automation Requests or produce flight plans that would violate the operating region constraints.
%
Some of the Line Search Task messages transmitted from the Ground Station contained malformed messages that would modify the UxAS behavior. In addition, we modified the Ground Station to simulate a breach for our evaluation of the attestation transform.
%
We performed a set of tests on the initial system (\figref{fig:sw-initial}) prior to applying our cyber-resiliency mitigations in order to verify the effectiveness of the malicious code.
A third party evaluator repeated the tests on the hardened system and demonstrated that our mitigation strategy was successful.
%
The following scenarios were exercised:

\paragraph{Infected Ground Station:\/} An application file is modified on the Ground Station, which sends the UAV on a mission-violating trajectory.
%
On the hardened system, messages sent to the UAV were rejected by the Attestation Manager.

\paragraph{Malformed Line Search Task message:\/}
%
The Line Search Task message contained a waypoint with a longitude outside the permitted range to exploit vulnerabilities in the on-board software.
%
The wellformedness filter for Line Search Task messages prevented this message from reaching the UxAS.

\paragraph{UxAS vulnerability exploit:\/} Line Search Tasks with greater than 90 waypoints were transmitted from the Ground Station, triggering a vulnerability that crashes UxAS.
%
When this occurs, UxAS is unable to generate an Automation Response.
%
The response monitor detected and signaled the Ground Station when UxAS failed to respond to the request.

\paragraph{UxAS trojan modifies flight plan:\/} A trojan embedded in UxAS attempts to cause the UAV to fly into a specified keep-out zone by modifying the mission command waypoints in the Automation Response.
%
On the hardened system the Geofence Monitor detected that it was being instructed to fly into a keep-out zone and returned the UAV to Home Base.

@@

The case study demonstrated the viability of writing code contracts and synthesizing implementations from code contracts in a real-world system.
%
It additionally provided insights for improvement.
%
We discovered very early that we needed a mechanism to test contracts.
%
We initially used the synthesis path with a custom test harness for the synthesized code, but this proved to be slow and cumbersome because we had to move back and forth between the OSATE environment and an external test platform.
%
Test contracts are a direct result of our experience, and they obviate the back and forth by integrating test into the modeling and analysis environment.

>>>>

The main organizing concept is that of an architecture-to-architecture \emph{security-improving transform}, achieved via the insertion of new architectural components aimed at mitigating cyber-vulnerabilities.
%
We describe two such transformations in this paper: (1) the insertion of a filter to prevent malformed data from a malicious actor being propagated to downstream components, and (2) the insertion of a runtime monitor to detect (and alert) unexpected behaviors arising from uncontrolled components.
%
We refer to these as \emph{high-assurance components} because they must be correct and fulfill their intended purpose for the system to be cyber-resilient.

@@

We report on a case study applying these transformations with \brfcs\ to an Unmanned Aerial Vehicle (UAV) system that uses the Air Force Research Laboratory's OpenUxAS services for route planning.
%
OpenUxAS, as an open source product, is considered \emph{uncontrolled}.
%
The UAV system is thus transformed to be resilient to malicious behavior that may arise in the uncontrolled component.
%
In particular, filters are added to guard against malformed input and a monitor is created to guard against malicious flight plans generated from OpenUxAS.
%
The case study system is complex and shows the viability of the approach in potential full-scale industrial design.

@@

Having passed \agr\ verification, the workflow in \figref{fig:workflow} proceeds to the thread analysis as the UAV must actively protect itself from malicious actors.
%
Threat analysis identifies the potential of a supply chain attack through the AI route planner since its third-party.
%
To mitigate the identified supply chain attack, the designer adds a requirement that the system must guard against malicious AI behavior.
%
This requirement is reflected in the component contracts by removing all output guarantees from the AI contract since that component is considered uncontrolled.
%
Without output guarantees, the AI component is able to output arbitrary events in the \agr\ verification analysis.

@@

As shown in \figref{fig:hardened-certificate}, the cyber-hardened system passes AGREE verification.
%
That means that the high-assurance components protect the WM from the uncontrolled AI generating malformed data, spontaneous responses, or delayed responses.
%
Synthesis must generate faithful implementations of the code contracts for the \agr\ verification result to apply to the deployed system.

@@

\item the predicate is evaluated on the input;

@@

In this section, we outline the application of the \brfcs\ tool towards the development of a UAV surveillance system.
%
The motivation was to showcase the utility and capability of the BriefCASE toolchain on an industrial scale problem.
%
Another part of motivation wast to determine whether or not the technology provides designers with early validation of their models,
%
produces early information about cyber-vulnerabilities,
%
provides effective mitigations,
%
and handles a complex real-world message format, namely the UxAS message format.

@@

For this application, UxAS accepts three types of messages.
%
The \textit{Operating Region} message defines where the UAV can and cannot fly.
%
The \textit{Line Search Task} message contains a series of waypoints that the UAV should traverse.
%
The waypoints typically lie along some geographical feature of interest, such as a river or railway.
%
Note that the UAV may not be able to directly traverse the Line Search Task waypoints due to constraints specified in the Operating Region message.
%
Anytime after receiving the Operating Region and Line Search Task messages, a Ground Station can transmit an \textit{Automation Request} message, which instructs UxAS to generate a flight plan that satisfies these constraints.
%
UxAS passes the flight plan in an \textit{Automation Response} message to the Waypoint Manager.
%
Because the Flight Control Computer can only process a small number of waypoints at a time, the Waypoint Manager parcels a small number of waypoints corresponding to the current UAV position, and sends them to the Flight Control Computer over a serial connection via the UART Driver.

@@

In \brfcs, we analyze the model using one (or more) of the integrated cybersecurity analysis tools.
%
As the designers of the system, we reviewed the list of potential vulnerabilities, and then added new requirements to the system model.
%
For example, the vulnerability analysis noted that UxAS is an open source project, so we marked that component as \texttt{uncontrolled} (colored red in \figref{fig:sw-initial}) and added requirements for ensuring that unverified or malicious code (which could potentially be embedded in the component) is not able impact other components.

In total, we added seven cyber requirements from the vulnerability analysis into our model.
%
These include four integrity requirements on data, two requirements for monitoring the behavior of the open-source UxAS component, and an attestation requirement for ensuring the integrity of the Ground Station software.
%
Requirements are imported into the model as goals in the assurance case.
%
Because we can build the assurance case at any time during development, we can easily determine for a given snapshot of the model which requirements are not yet supported by evidence.

The intent of the data integrity requirements is to prevent malformed messages from causing a buffer overrun or code injection attack.
%
In the UAV design, such messages are most likely to originate from a remote source or the uncontrolled UxAS component.
%
By placing filters on the connections upstream of mission-critical components, such attacks could be mitigated.
%
We therefore inserted filter components on the incoming and outgoing UxAS connections.

The filter code contracts for the four UxAS connections are similar in that they check that record values contained in the messages are within appropriate ranges.
%
For example, the Automation Response message filter, which drops messages containing malformed flight plans, is defined as shown in \figref{fig:automation-response-filter}.
%
Although \texttt{Latitude}, \texttt{Longitude} and \texttt{Altitude} are defined as 64-bit floating-point values, the filter only passes messages containing waypoint values between [-90,90], [-180,180], and [0,15000], respectively.

@@

In addition to monitoring the UxAS output for malformed messages, we must also monitor for suspicious behavior.
%
This requires adding components for detecting that UxAS has crashed, as well as monitoring the correctness of the flight plans it produces.
%
We therefore add a response monitor (Resp\_Monitor) to send an alert if UxAS does not emit a response within a set amount of time from receiving a request.
%
And we add a geo-fence monitor (Geofence\_Monitor) to ensure that generated flight plans are compliant with the specified keep-in and keep-out zones.
%
The Geofence\_Monitor is a gated monitor in that it prevents the observed automation response from reaching the Waypoint Manager if the zones are violated.
%
Its code contract is shown in \figref{fig:geofence-monitor}.

@@

So far, we have addressed requirements that mitigate vulnerabilities related to malformed messages and malicious behavior on-board the UAV.
%
But we also want to protect against a compromised ground station transmitting well-formed, but malicious, commands.
%
As such we added two components to the UAV software: the Attestation Manager for evaluating remote systems like the Ground Station, and the Attestation Gate for filtering messages from sources that have not been approved by the Attestation Manager.
%
The Attestation Manager \ckml\ implementation is automatically inserted into the application code base by \brfcs.
%
The code contract for the Attestation Gate is generated by \brfcs\ since its behavior is specific to attestation.

After transforming the model to address the cyber requirements, the software architecture now appears as shown in \figref{fig:hardened-sw}.
%
The shaded components (green) were added by \brfcs\ to harden the system with the exception of the UxAS component (red).

After unit testing the monitors, we formally verified the model with \agr\ to show that all of the component contracts were satisfied.
%
We then ran \splt\ to produce provably-correct code.

@@

The demonstration system was then assembled for testing.
%
\brfcs\ includes the HAMR tool to generate the component stubs and infrastructure code necessary to enable component communication and execution according to a specified schedule.
%
HAMR was used here to target the verified micro-kernel seL4.
%
The UAV and Ground Station software were deployed on ODROID XU4 hardware, and communicated with each other over ethernet.
%
The AMASE flight simulator, representing the Flight Control Computer was run on a Linux machine and connected to the UAV ODROID via a serial connection.

The UxAS implementation on the UAV was modified by adding malicious code that would prevent it from responding to Automation Requests or produce flight plans that would violate the constraints in the Operating Region message.
%
Some of the Line Search Task messages transmitted from the Ground Station contained malformed messages that would modify the UxAS behavior. In addition, we modified the Ground Station to simulate a breach for our evaluation of the attestation.

We performed a set of tests on the initial system (\figref{fig:sw-initial}) prior to applying our cyber-resiliency mitigation in order to verify the effectiveness of the malicious code.
A third party evaluator repeated the tests on the hardened system and demonstrated that our mitigation strategy was successful.
The following scenarios were exercised:

\paragraph{Infected Ground Station:\/} an application file is modified on the Ground Station, which sends the UAV on a mission-violating trajectory.
%
On the hardened system, messages sent to the UAV were rejected by the Attestation Manager.

\paragraph{Malformed Line Search Task message:\/}
%
the Line Search Task message contained a waypoint with a longitude outside the permitted range to exploit vulnerabilities in the on-board software.
%
The filter for Line Search Task messages prevented this message from reaching the UxAS.

\paragraph{UxAS vulnerability exploit:\/} tasks with greater than 90 waypoints were transmitted from the Ground Station, triggering a vulnerability that crashes UxAS.
%
When this occurs, UxAS is unable to generate an Automation Response.
%
The response monitor detected and signaled the Ground Station when UxAS failed to respond to the request.

\paragraph{UxAS trojan modifies flight plan:\/} a trojan embedded in UxAS attempts to cause the UAV to fly into a specified keep-out zone by modifying the mission command waypoints in the Automation Response.
%
On the hardened system the Geofence Monitor detected that it was being instructed to fly into a keep-out zone and returned the UAV to Home Base.

@@

The case study demonstrated the viability of writing code contracts and synthesizing implementations from code contracts in a real-world system.
%
It additionally provided insights for improvement.
%
We discovered very early that we needed a mechanism to test contracts.
%
We initially used the synthesis path with a custom test harness for the synthesized code, but this proved to be slow and cumbersome because we had to move back and forth between the \brfcs\ environment and an external test platform.
%
Test contracts are a direct result of our experience, and they obviate the back and forth by integrating test into the modeling and analysis environment.
